---
title: "SQL by Connor Keeling"
output: html_document
---
## Filtering Rows and Columns

In SQL, you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First, we should load the Lahman package.We will also need the sqldf package. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example.Suppose we want to see the home run totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Let's find all instances where Yankee players have hit 40 or more homeruns in a season:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Now let's find all instances where a MLB player had over 40 homeruns but fewer than 60 strikeouts:

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO from Batting WHERE HR>40 and SO<60"
sqldf(query)
```

Finally, let's find all the instances of Phillies players who hit more than 30 homeruns in a season in the 1970's:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting WHERE HR>30 and yearID>=1970 and yearID<=1979 and teamID='PHI'"
sqldf(query)
```

## ORDER BY

With SQL we can also order the rows in the output. For example, suppose you want to see every instance of a player hitting over 50 homeruns, but you want to see the hitters with the most homeruns appear at the top. We can do that like this:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting WHERE
HR>50
ORDER BY HR DESC"
sqldf(query)
```

## AGGREGATION

We can also squash figures together using aggregation.  Say we want to find the career total homeruns for Babe Ruth. Rather than adding the HR totals from each individual season, we can do this:

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

Let's say we want to find the career homerun totals for all baseball players, but that we only want to list those with 600 or more HR.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)

```

We can use this method to find the average totals of homeruns, as well. Let's find out which players have the highest per-season HR averages.

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting 
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)

```

## JOINS

We can inner join the data to fully list the information.  Let's say we want to list Babe Ruth's full name along with his per-season homerun totals:

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Now let's find all instances of players with over 50 homeruns in a season while listing first and last names, year, team name, and number of homeruns:

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
Next, let's list Babe Ruth with his playerID, yearID, team name, and number of homeruns:

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```

Next, let's list Babe Ruth again
```{r}
colnames(Master)
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Let's find carer homerun totals for players but limit the display to those with 600 or more homeruns and list them from most to fewest homeruns:

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR) >=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

Finally, let's find all players with the first name Bob who averaged over $1 million in salary in their career, and list their last names:

```{r}
query<-"SELECT nameFirst,nameLast,avg(Salary)
FROM Salaries INNER JOIN Master 
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(Salary)>=1000000
ORDER BY avg(Salary) DESC"

sqldf(query)
```

